### 2025-03-01 README.Rmd
transfer repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# publishTC

Why not publish the trend-cycle?
The goal of `publishTC` is to facilitate the computation of trend-cycle component:

- Using the Cascade Linear Filter (CLF) and the surrogate cut-and-normalise asymmetric filters, as done by [Statistique Canada](https://www.statcan.gc.ca/en/dai/btd/trend-cycle);

- Using the classical Henderson symmetric filter and the surrogate Musgrave asymmetric filters, as done by Australian Bureau of Statistics (Trewin 2003);

- Using a local Parametrization of the Musgrave asymmetric filters, as described in Quartier-la-Tente (2024);

- Extending the Henderson symmetric fiter and the surrogate Musgrave asymmetric filters to take into account additive outliers and level shifts, as described in Quartier-la-Tente (2025).

# Bibliography

Dagum, E. B., & Luati, A. (2008). A Cascade Linear Filter to Reduce Revisions and False Turning Points for Real Time Trend-Cycle Estimation. *Econometric Reviews* 28 (1-3): 40‑59. <https://doi.org/10.1080/07474930802387837>

Henderson, R. (1916). Note on graduation by adjusted average. *Transactions of the actuarial society of America* 17: 43‑48.

Musgrave, J. (1964). A set of end weights to end all end weights. *US Census Bureau [custodian]*.

Trewin, D. (2003). A guide to interpreting time series - Monitoring trends. *Australian Bureau of Statistics Information Paper*. <https://www.abs.gov.au/AUSSTATS/abs@.nsf/Lookup/1349.0Main+Features12003?OpenDocument>.

Quartier-la-Tente, A. (2024). Improving Real-Time Trend Estimates Using Local Parametrization of Polynomial Regression Filters. *Journal of Official Statistics, 40*(4), 685-715. <https://doi.org/10.1177/0282423X241283207>.

Quartier-la-Tente, A. (2025). Estimation de la tendance-cycle avec des méthodes robustes aux points atypiques. <https://github.com/AQLT/robustMA>.