### 2025-03-01 report/CONJ_COMD.qmd
template creation

---
format: 
  html:
    toc: true
    number-sections: true
    embed-ressources: true
    number-depth: 2
date: today
date-format: long
params:
  file: "../data/CONJ_COMD.yml"
execute:
  cache: true
author:
  - name: Alain Quartier-la-Tente 
    affiliation: LEMNA
    orcid: 0000-0001-7890-3857
comments: 
    hypothesis:
      theme: clean
echo: false
eval: true
---

```{r}
#| include: false
##| cache: false
library(publishTC)
library(ggplot2)
library(patchwork)
spec <- yaml::read_yaml(params$file)
spec_series <- spec$series
list_files <- list.files(pattern = "csv$", path = file.path(dirname(params$file), spec$dataset), full.names = TRUE, include.dirs = FALSE)
all_data <- lapply(list_files, read.ts, list = TRUE)
names(all_data) <- tools::file_path_sans_ext(basename(list_files))
n <- length(all_data)
series <- names(all_data[[1]])
names(series) <- series

studied_methods <- sapply(spec$methods, `[[`, "eval")
names_methods <- sapply(spec$methods, `[[`, "name")[which(studied_methods)]
studied_methods <- names(studied_methods)[which(studied_methods)]
names(studied_methods) <- names_methods
digits <- spec$plots$digits

last_est <- lapply(series, function(serie) {
	spec_s <- spec$series[[serie]]
	x <- window(all_data[[n]][[serie]], start = spec_s$first_date)
	smoothing(x = x, methods = studied_methods, length = spec_s$length, 
			  ao = spec_s$outliers$ao,
			  ao_tc = spec_s$outliers$ao_tc,
			  ls = spec_s$outliers$ls)
})

if (spec$plots$extra_info$enabled) {
	extra_info <- lapply(last_est, function(sect) {
		res <- lapply(sect, summary)
		res <- do.call(
			rbind,
			lapply(res, function(x){
				x[["I/C ratios"]] <- NULL
				unlist(x)
			}))
		rownames(res) <- gsub("\n", " ", rownames(res))
		res
	})

	if (length(grep("robust", studied_methods)) > 0) {
		outliers <- lapply(spec_series, `[[`, "outliers")
	} else {
		outliers <- lapply(spec_series, function(x) NULL)
	}

}

if (spec$plots$revision_history$enabled) {
	rev_to_table <- function(x) {
		data_rev <- t(do.call(ts.union, x))
		data_rev <- data_rev[,ncol(data_rev):1] # last date first
		data_rev <- data.frame(
			`Last date` = gsub("_", "-", rep(colnames(x[[1]]), length(x))),
			Methods = rep(names(x), each = ncol(x[[1]])), 
			data_rev,
			check.names = FALSE)
		colnames(data_rev)[-(1:2)] <- rev(format(zoo::as.Date(time(x[[1]])), "%Y-%m"))
		rownames(data_rev) <- NULL
		data_rev
	}
	dt_table <- function(data, group_cols = 2, i_row_name = 1, digits = spec$plots$revision_history$table$digits,
						 caption = NULL) {
		group_cols[group_cols > i_row_name] <- group_cols[group_cols > i_row_name] - 1
		is_num <- which(sapply(data[,-i_row_name], is.numeric))
		DT::datatable(
			data[,-i_row_name],
			extensions = c("RowGroup", "Buttons"),
			options = list(
				rowGroup = list(dataSrc = group_cols),
				columnDefs = list(list(targets = group_cols, visible = FALSE)),
				dom = "Brtipl",
				buttons = c("copy", "csv", "excel", "pdf")
			),
			selection = "none",
			rownames = data[,i_row_name],
			caption = caption
		) |>  DT::formatRound(columns = is_num,
							  digits= digits)
	}

	rev_data <- lapply(basename(list_files), function(f){
		res <- lapply(studied_methods, function(method) {
			read.ts(file.path(dirname(params$file), spec$dataset, method, f), list = TRUE)
		})
		c(list("Raw data" = lapply(all_data[[tools::file_path_sans_ext(f)]], window, 
								   start = min(sapply(spec_series, `[[`, "first_date")))),
		  res)
	})
	names(rev_data) <- names(all_data)
	rev_data <- rev(rev_data)
	rev_data_ <- lapply(names(rev_data[[1]][[1]]), function(serie){
		res <- lapply(names(rev_data[[1]]), function(method){
			do.call(ts.union, lapply(rev_data, function(x){
				x[[method]][[serie]]
			}))
		})
		names(res) <- names(rev_data[[1]])
		res
	})
	names(rev_data_) <- names(rev_data[[1]][[1]])
	rev_data <- rev_data_
	rm(rev_data_)

	if (spec$plots$revision_history$table$contribution_of_sa) {
		rev_data_fixed <- lapply(series, function(serie) {
			spec_s <- spec$series[[serie]]
			y <- window(all_data[[n]][[serie]], start = spec$series[[serie]]$first_date)
			vintages <- lapply(all_data[-n], function(vintage){
				window(y, end = end(vintage[[serie]]))
			})
			lapply(studied_methods, function(method) {
				method_name <- names(studied_methods)[which(studied_methods == method)]
				res <- c(lapply(vintages, function(x) {

					smoothing(x = x, methods = method, length = spec_s$length, 
							  ao = spec_s$outliers$ao,
							  ao_tc = spec_s$outliers$ao_tc,
							  ls = spec_s$outliers$ls)[[1]][["tc"]]
				}),
				list(last_est[[serie]][[method_name]][["tc"]])
				)
				res <- do.call(ts.union, rev(res))
				colnames(res) <- colnames(rev_data[[serie]][[method_name]])
				res
			})
		})

		rev_tot <- lapply(series, function(serie) {
			rev_by_method <- lapply(names(studied_methods), function(method){
				data <- rev_data_fixed[[serie]][[method]]
				data_tot <- rev_data[[serie]][[method]]
				rev_tc <- (data[,1] - data)
				rev_tot <- (data_tot[,1] - data_tot)
				rev_tot <- rev_tot
				rev_sa <- rev_tot - rev_tc
				colnames(rev_tc) <- colnames(rev_sa) <- colnames(rev_tot) <- colnames(data_tot)
				list(
					Total = rev_tot, 
					`Due to revisions on SA` = rev_sa, 
					`Due to update of TC model` = rev_tc
				)
			})
			names(rev_by_method) <- names(studied_methods)
			rev_tot_by_method <- do.call(
				rbind,
				lapply(names(rev_by_method), function(method){
					rev_meth <- rev_to_table(rev_by_method[[method]])
					cbind(Meth = method, rev_meth)
				})
			)
			rev_tot_by_method <- rev_tot_by_method[rev_tot_by_method[, "Last date"] != colnames(rev_tot_by_method)[4], -4]
			rev_tot_by_method
		})
	}

}

plots <- unlist(spec$plots$kind)
plots <- names(plots)[which(plots)]
nrow_legend <- sum(c(any(plots %in% c("normal", "lollypop")), any(plots %in% c("implicit_forecast", "confint"))))
```

---
title: "`r spec$datasetname`"
---

```{r}
#| results: asis

for (serie in names(all_data[[1]])) {
	last_est_s <- last_est[[serie]]

	all_plots <- lapply(last_est_s, function(est) {
		plots_ <- ggsmoothing_plot(est, plots = plots)
		if (any(c("implicit_forecast", "confint") %in% plots)) {
			if ("normal" %in% plots) {
				plots_$normal <- plots_$normal + theme(legend.position = element_blank())
			}
			if ("lollypop" %in% plots) {
				plots_$lollypop <- plots_$lollypop + theme(legend.position = element_blank())
			}
		}
		plots_
	})

	cat("# ", spec_series[[serie]]$description, "\n")
	if (spec$plots$bymethods$enabled){
		cat("## ", spec$plots$bymethods$name)
		cat("\n\n::::: {.panel-tabset}\n")
		for (methods in names(last_est_s)) {
			cat(sprintf("\n### %s\n\n", gsub("\n", " ", methods)))
			first_date <- time(last_est_s[[methods]]$tc)[length(last_est_s[[methods]]$tc)] - spec$plots$nyears
			range_y <- range(window(ts.union(last_est_s[[methods]]$tc, last_est_s[[methods]]$x), start = first_date))
			cat(knitr::knit_child(text = c(
				"```{r}",
				"#| fig-height: 8",
				"#| fig-width: 8",
				'wrap_plots(all_plots[[methods]]) +
			plot_layout(axes = "collect") &
			theme_bw() & theme(legend.title = element_blank(), legend.position = "bottom") &
			# guides(color=guide_legend(nrow=nrow_legend)) &
			labs(x = NULL, y = NULL) &
			coord_cartesian(xlim = c(first_date, NA), y = range_y)',
			"```",
			""
			), envir = environment(), quiet = TRUE
			))
		}
		cat("\n:::::\n\n")
	}

	if (spec$plots$byplots$enabled){
		byplots <- lapply(plots, function(kind) {
			res <- lapply(names(all_plots), function(method) {
				all_plots[[method]][[kind]] + ggtitle(method)
			})
			names(res) <- names(all_plots)
			res
		})
		names(byplots) <- plots

		cat("## ", spec$plots$byplots$name)
		cat("\n\n::::: {.panel-tabset}\n")
		for (plot in names(byplots)) {
			cat(sprintf("\n### %s\n\n", plot))
			first_date <- time(last_est_s[[1]]$tc)[length(last_est_s[[1]]$tc)] - spec$plots$nyears
			range_y <- range(window(ts.union(last_est_s[[1]]$tc, last_est_s[[1]]$x), start = first_date))
			cat(knitr::knit_child(text = c(
				"```{r}",
				"#| fig-height: 8",
				"#| fig-width: 8",
				'wrap_plots(byplots[[plot]]) +
			plot_layout(guides = "collect", axes = "collect") &
			theme_bw() & theme(legend.title = element_blank(), legend.position = "bottom") &
			labs(x = NULL, y = NULL) &
			coord_cartesian(xlim = c(first_date, NA), y = range_y) ',
			"```",
			""
			), envir = environment(), quiet = TRUE
			))
		}

		cat("\n:::::\n\n")
	}

	if (spec$plots$comparison_plot$enabled) {

		cat("## ", spec$plots$comparison_plot$name)

		data_comp <- do.call(ts.union, lapply(last_est_s, `[[`, "tc"))
		first_date <- time(last_est_s[[1]]$tc)[length(last_est_s[[1]]$tc)] - spec$plots$nyears
		range_y <- range(unlist(lapply(last_est_s, function(x) range(window(ts.union(x$tc), start = first_date)))))
		cat("\n\n::::: {.panel-tabset}\n")

		cat("\n### Short\n")
		cat(knitr::knit_child(text = c(
			"```{r}",
			"#| fig-height: 8",
			"#| fig-width: 8",
			'autoplot(data_comp) +
			theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom") +
			labs(x = NULL, y = NULL) +
			coord_cartesian(xlim = c(first_date, NA), y = range_y) ',
			"```",
			""
		), envir = environment(), quiet = TRUE
		))
		cat("\n### All\n")

		cat(knitr::knit_child(text = c(
			"```{r}",
			"#| fig-height: 8",
			"#| fig-width: 8",
			'p <- autoplot(data_comp) +
	theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom") +
	labs(x = NULL, y = NULL)',
	ifelse(spec$plots$comparison_plot$interactive, 'plotly::ggplotly(p)', 'p'),
	"```",
	""
		), envir = environment(), quiet = TRUE
		))

		cat("\n:::::\n")
	}

	if (spec$plots$revision_history$enabled) {
		rev_data_s <- rev_data[[serie]]

		cat("## ", spec$plots$revision_history$name)

		cat("\n\n::::: {.panel-tabset}\n")
		for (methods in names(rev_data_s)) {
			cat(sprintf("\n### %s\n\n", gsub("\n", " ", methods)))
			first_date <- time(rev_data_s[[methods]])[length(time(rev_data_s[[methods]]))] - spec$plots$nyears
			range_y <- range(window(rev_data_s[[methods]], start = first_date), na.rm = TRUE)
			cat(knitr::knit_child(text = c(
				"```{r}",
				"#| fig-height: 8",
				"#| fig-width: 8",
				'autoplot(round(rev_data_s[[methods]], digits)) +
			theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom") +
			labs(x = NULL, y = NULL) +
			coord_cartesian(xlim = c(first_date, NA), y = range_y)',
			"```",
			""
			), envir = environment(), quiet = TRUE
			))
		}

		cat("\n:::::\n")

		if (spec$plots$revision_history$table$enabled) {
			cat("## ", spec$plots$revision_history$name)
			cat("\n### Revisions\n")
			data_rev <- rev_to_table(rev_data_s)
			cat(knitr::knit_child(text = c(
				"```{r}",
				'dt_table(data_rev)',
				"```",
				""
			), envir = environment(), quiet = TRUE
			))
			if (spec$plots$revision_history$table$contribution_of_sa) {
				cat("\n### Revisions to last estimates\n")
				cat(knitr::knit_child(text = c(
					"```{r}",
					'dt_table(rev_tot[[serie]], group_cols = 1:2, i_row_name = 3)',
					"```",
					""
				), envir = environment(), quiet = TRUE
				))
			}

		}
	}

	if (spec$plots$extra_info$enabled) {
		cat("## ", spec$plots$extra_info$name)

		cat(knitr::knit_child(text = c(
			"```{r}",
			'knitr::kable(extra_info[[serie]],align = "lccc", digits = digits)',
			"```",
			""
		), envir = environment(), quiet = TRUE
		))
		if (any(lengths(outliers[[serie]]) > 0)) {
			cat("**Outliers**:\n\n")
			if (!is.null(outliers[[serie]]$ao)) {
				cat(
					"\n\n- *Additive outlier* (AO) :",
					knitr::combine_words(format(zoo::as.yearmon(outliers[[serie]]$ao), "%B %Y")),
					"\n\n"
				)
			}

			if (!is.null(outliers[[serie]]$ao_tc)) {
				cat(
					"\n\n- - *Additive outlier* (AO) associated to Trend-Cycle:",
					knitr::combine_words(format(zoo::as.yearmon(outliers[[serie]]$ao), "%B %Y")),
					"\n\n"
				)
			}

			if (!is.null(outliers[[serie]]$ls)) {
				cat(
					"\n\n- *Level shift* (LS):",
					knitr::combine_words(format(zoo::as.yearmon(outliers[[serie]]$ao), "%B %Y")),
					"\n\n"
				)
			}

		}

	}
}
```