### 2025-03-20 report/CONJ_COMD.qmd
update template

---
format: 
  html:
    toc: true
    fig-format: png
    number-sections: true
    embed-resources: true
    number-depth: 2
    other-links:
      - text: More surveys/data
        icon: patch-plus
        href: https://aqlt.github.io/publishTC.wp/
    code-links:
      - text: AQLT/publishTC.wp
        icon: github
        href: https://github.com/AQLT/publishTC.wp
date: today
date-format: long
params:
  file: "../data/CONJ_COMD.yml"
# execute:
#   cache: true
author:
  - name: Alain Quartier-la-Tente 
    affiliation: LEMNA
    orcid: 0000-0001-7890-3857
echo: false
eval: true
fig-height: 6
fig-width: 8
---

```{r}
#| include: false
#| cache: false
library(publishTC)
library(ggplot2)
library(patchwork)
library(plotly)
spec <- yaml::read_yaml(params$file)
spec_series <- spec$series
series <- names(spec_series)
names(series) <- series
list_files <- list.files(pattern = "csv$", path = file.path(dirname(params$file), spec$dataset), full.names = TRUE, include.dirs = FALSE)
all_data <- lapply(list_files, read.ts, list = TRUE)
names(all_data) <- tools::file_path_sans_ext(basename(list_files))
all_data <- lapply(all_data, `[`, series)
n <- length(all_data)
if (n == 1) {
    spec$plots$revision_history$enabled <- FALSE
}

studied_methods <- sapply(spec$methods, `[[`, "eval")
names_methods <- sapply(spec$methods, `[[`, "name")[which(studied_methods)]
studied_methods <- names(studied_methods)[which(studied_methods)]
names(studied_methods) <- names_methods
digits <- spec$plots$digits

last_est <- lapply(series, function(serie) {
	spec_s <- spec$series[[serie]]
	x <- window(all_data[[n]][[serie]], start = spec_s$first_date)
	smoothing(x = x, methods = studied_methods, length = spec_s$length, 
			  ao = spec_s$outliers$ao,
			  ao_tc = spec_s$outliers$ao_tc,
			  ls = spec_s$outliers$ls)
})

plots <- unlist(spec$plots$kind)
plots <- names(plots)[which(plots)]
nrow_legend <- sum(c(any(plots %in% c("normal", "lollypop")), any(plots %in% c("implicit_forecast", "confint"))))
```

---
title: "`r spec$datasetname`"
---

```{r}
#| label: extra-info
#| eval: !expr spec$plots$extra_info$enabled
extra_info <- lapply(last_est, function(sect) {
    res <- lapply(sect, summary)
    res <- do.call(
        rbind,
        lapply(res, function(x){
            x[["I/C ratios"]] <- NULL
            unlist(x)
        }))
    rownames(res) <- gsub("\n", " ", rownames(res))
    res
})

if (length(grep("robust", studied_methods)) > 0) {
    outliers <- lapply(spec_series, `[[`, "outliers")
} else {
    outliers <- lapply(spec_series, function(x) NULL)
}
```

```{r}
#| label: revision-history
#| eval: !expr spec$plots$revision_history$enabled
rev_to_table <- function(x) {
    data_rev <- t(do.call(ts.union, x))
    data_rev <- data_rev[,ncol(data_rev):1] # last date first
    data_rev <- data.frame(
        `Last date` = gsub("_", "-", rep(colnames(x[[1]]), length(x))),
        Methods = rep(names(x), each = ncol(x[[1]])), 
        data_rev,
        check.names = FALSE)
    colnames(data_rev)[-(1:2)] <- rev(format(zoo::as.Date(time(x[[1]])), "%Y-%m"))
    rownames(data_rev) <- NULL
    data_rev
}
dt_table <- function(data, group_cols = 2, i_row_name = 1, nb_digits = digits,
                     caption = NULL) {
    group_cols[group_cols > i_row_name] <- group_cols[group_cols > i_row_name] - 1
    is_num <- which(sapply(data[,-i_row_name], is.numeric))
    DT::datatable(
        data[,-i_row_name],
        extensions = c("RowGroup", "Buttons"),
        options = list(
            rowGroup = list(dataSrc = group_cols),
            columnDefs = list(list(targets = group_cols, visible = FALSE)),
            dom = "Brtipl",
            buttons = c("copy", "csv", "excel", "pdf")
        ),
        selection = "none",
        rownames = data[,i_row_name],
        caption = caption
    ) |>  DT::formatRound(columns = is_num,
                          digits= nb_digits)
}

rev_data <- lapply(basename(list_files), function(f){
    res <- lapply(studied_methods, function(method) {
        read.ts(file.path(dirname(params$file), spec$dataset, method, f), list = TRUE)[series]
    })
    c(list("Raw data" = lapply(all_data[[tools::file_path_sans_ext(f)]], window, 
                               start = min(sapply(spec_series, `[[`, "first_date")))),
      res)
})
names(rev_data) <- names(all_data)
rev_data <- rev(rev_data)
rev_data_ <- lapply(names(rev_data[[1]][[1]]), function(serie){
    res <- lapply(names(rev_data[[1]]), function(method){
        do.call(ts.union, lapply(rev_data, function(x){
            x[[method]][[serie]]
        }))
    })
    names(res) <- names(rev_data[[1]])
    res
})
names(rev_data_) <- names(rev_data[[1]][[1]])
rev_data <- rev_data_
rm(rev_data_)
```

```{r}
#| label: revision-history-contribution-sa
#| eval: !expr spec$plots$revision_history$enabled & spec$plots$revision_history$table$contribution_of_sa
rev_data_fixed <- lapply(series, function(serie) {
    spec_s <- spec$series[[serie]]
    y <- window(all_data[[n]][[serie]], start = spec$series[[serie]]$first_date)
    vintages <- lapply(all_data[-n], function(vintage){
        window(y, end = end(vintage[[serie]]))
    })
    lapply(studied_methods, function(method) {
        method_name <- names(studied_methods)[which(studied_methods == method)]
        res <- c(lapply(vintages, function(x) {

            smoothing(x = x, methods = method, length = spec_s$length, 
                      ao = spec_s$outliers$ao,
                      ao_tc = spec_s$outliers$ao_tc,
                      ls = spec_s$outliers$ls)[[1]][["tc"]]
        }),
        list(last_est[[serie]][[method_name]][["tc"]])
        )
        res <- do.call(ts.union, rev(res))
        colnames(res) <- colnames(rev_data[[serie]][[method_name]])
        res
    })
})

rev_tot <- lapply(series, function(serie) {
    rev_by_method <- lapply(names(studied_methods), function(method){
        data <- rev_data_fixed[[serie]][[method]]
        data_tot <- rev_data[[serie]][[method]]
        rev_tc <- (data[,1] - data)
        rev_tot <- (data_tot[,1] - data_tot)
        rev_tot <- rev_tot
        rev_sa <- rev_tot - rev_tc
        colnames(rev_tc) <- colnames(rev_sa) <- colnames(rev_tot) <- colnames(data_tot)
        list(
            Total = rev_tot, 
            `Due to revisions on SA` = rev_sa, 
            `Due to update of TC model` = rev_tc
        )
    })
    names(rev_by_method) <- names(studied_methods)
    rev_tot_by_method <- do.call(
        rbind,
        lapply(names(rev_by_method), function(method){
            rev_meth <- rev_to_table(rev_by_method[[method]])
            cbind(Meth = method, rev_meth)
        })
    )
    rev_tot_by_method <- rev_tot_by_method[rev_tot_by_method[, "Last date"] != colnames(rev_tot_by_method)[4], -4]
    rev_tot_by_method[with(rev_tot_by_method, order(Meth,`Last date`, Methods)), ]
})
```

```{r}
#| results: asis

for (serie in names(all_data[[1]])) {
	last_est_s <- last_est[[serie]]

	all_plots <- lapply(last_est_s, function(est) {
		plots_ <- ggsmoothing_plot(est, plots = plots)
		if (any(c("implicit_forecast", "confint") %in% plots)) {
			if ("normal" %in% plots) {
				plots_$normal <- plots_$normal + theme(legend.position = element_blank())
			}
			if ("lollypop" %in% plots) {
				plots_$lollypop <- plots_$lollypop + theme(legend.position = element_blank())
			}
		}
		plots_
	})

	cat("# ", spec_series[[serie]]$description, "\n")
	if (spec$plots$bymethods$enabled){
		cat("## ", spec$plots$bymethods$name)
		cat("\n\n::::: {.panel-tabset}\n")
		for (i in seq_along(last_est_s)) {
		    methods <- names(last_est_s)[i]
			cat(sprintf("\n### %s\n\n", gsub("\n", " ", methods)))
			first_date <- time(last_est_s[[methods]]$tc)[length(last_est_s[[methods]]$tc)] - spec$plots$nyears
			range_y <- range(window(ts.union(last_est_s[[methods]]$tc, last_est_s[[methods]]$x), start = first_date))
			cat(knitr::knit_child(text = c(
				"```{r}",
				sprintf("#| label: bymethods-%s-%s", serie, studied_methods[i]),
				'wrap_plots(all_plots[[methods]]) +
			plot_layout(axes = "collect") &
			theme_bw() & theme(legend.title = element_blank(), legend.position = "bottom") &
			# guides(color=guide_legend(nrow=nrow_legend)) &
			labs(x = NULL, y = NULL) &
			coord_cartesian(xlim = c(first_date, NA), y = range_y)',
			"```",
			""
			), envir = environment(), quiet = TRUE
			))
		}
		cat("\n:::::\n\n")
	}

	if (spec$plots$byplots$enabled){
		byplots <- lapply(plots, function(kind) {
			res <- lapply(names(all_plots), function(method) {
				all_plots[[method]][[kind]] + ggtitle(method)
			})
			names(res) <- names(all_plots)
			res
		})
		names(byplots) <- plots

		cat("## ", spec$plots$byplots$name)
		cat("\n\n::::: {.panel-tabset}\n")
		for (plot in names(byplots)) {
			cat(sprintf("\n### %s\n\n", plot))
			first_date <- time(last_est_s[[1]]$tc)[length(last_est_s[[1]]$tc)] - spec$plots$nyears
			range_y <- range(window(ts.union(last_est_s[[1]]$tc, last_est_s[[1]]$x), start = first_date))
			cat(knitr::knit_child(text = c(
				"```{r}",
				sprintf("#| label: byplots-%s-%s", serie, plot),
				'wrap_plots(byplots[[plot]]) +
			plot_layout(guides = "collect", axes = "collect") &
			theme_bw() & theme(legend.title = element_blank(), legend.position = "bottom") &
			labs(x = NULL, y = NULL) &
			coord_cartesian(xlim = c(first_date, NA), y = range_y) ',
			"```",
			""
			), envir = environment(), quiet = TRUE
			))
		}

		cat("\n:::::\n\n")
	}

	if (spec$plots$comparison_plot$enabled) {

		cat("## ", spec$plots$comparison_plot$name)

		data_comp <- do.call(ts.union, c(list("y" = last_est_s[[1]][["x"]]), lapply(last_est_s, `[[`, "tc")))
		first_date <- time(data_comp)[nrow(data_comp)] - spec$plots$nyears
		range_y <- range(window(data_comp, start = first_date))
		cat("\n\n::::: {.panel-tabset}\n")

		cat("\n### Short\n")
		cat(knitr::knit_child(text = c(
			"```{r}",
				sprintf("#| label: comparison-%s", serie),
			'autoplot(data_comp) +
			theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom") +
			labs(x = NULL, y = NULL) +
			coord_cartesian(xlim = c(first_date, NA), y = range_y) ',
			"```",
			""
		), envir = environment(), quiet = TRUE
		))
		cat("\n### All\n")

		cat(knitr::knit_child(text = c(
			"```{r}",
			sprintf("#| label: comparisonfull-%s", serie),
			'p <- autoplot(data_comp) +
	theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom") +
	labs(x = NULL, y = NULL)',
	ifelse(spec$plots$comparison_plot$interactive, 'plotly::ggplotly(p)', 'p'),
	"```",
	""
		), envir = environment(), quiet = TRUE
		))

		cat("\n:::::\n")
	}

	if (spec$plots$revision_history$enabled) {
		rev_data_s <- rev_data[[serie]]

		cat("## ", spec$plots$revision_history$name)

		cat("\n\n::::: {.panel-tabset}\n")
		for (i in seq_along(rev_data_s)) {
		    methods <- names(rev_data_s)[i]
			cat(sprintf("\n### %s\n\n", gsub("\n", " ", methods)))
			first_date <- time(rev_data_s[[methods]])[length(time(rev_data_s[[methods]]))] - spec$plots$nyears
			range_y <- range(window(rev_data_s[[methods]], start = first_date), na.rm = TRUE)
			cat(knitr::knit_child(text = c(
				"```{r}",
				"#| fig-height: 5",
				sprintf("#| label: revisions-%s-%s", serie, c("rav", studied_methods)[i]),
				'autoplot(round(rev_data_s[[methods]], digits)) +
			theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom") +
			labs(x = NULL, y = NULL) +
			coord_cartesian(xlim = c(first_date, NA), y = range_y)',
			"```",
			""
			), envir = environment(), quiet = TRUE
			))
		}

		cat("\n:::::\n")

		if (spec$plots$revision_history$table$enabled) {
			cat("## ", spec$plots$revision_history$name)
			cat("\n### Revisions\n")
			data_rev <- rev_to_table(rev_data_s)
			cat(knitr::knit_child(text = c(
				"```{r}",
				'dt_table(data_rev)',
				"```",
				""
			), envir = environment(), quiet = TRUE
			))
			if (spec$plots$revision_history$table$contribution_of_sa) {
				cat("\n### Revisions to last estimates\n")
				cat(knitr::knit_child(text = c(
					"```{r}",
					'dt_table(rev_tot[[serie]], group_cols = 1:2, i_row_name = 3)',
					"```",
					""
				), envir = environment(), quiet = TRUE
				))
			}

		}
	}

	if (spec$plots$extra_info$enabled) {
		cat("## ", spec$plots$extra_info$name)

		cat(knitr::knit_child(text = c(
			"```{r}",
			'knitr::kable(extra_info[[serie]],align = "lccc", digits = digits)',
			"```",
			""
		), envir = environment(), quiet = TRUE
		))
		if (any(lengths(outliers[[serie]]) > 0)) {
			cat("**Outliers**:\n\n")
			if (!is.null(outliers[[serie]]$ao)) {
				cat(
					"\n\n- *Additive outlier* (AO) :",
					knitr::combine_words(format(zoo::as.yearmon(outliers[[serie]]$ao), "%B %Y")),
					"\n\n"
				)
			}

			if (!is.null(outliers[[serie]]$ao_tc)) {
				cat(
					"\n\n- - *Additive outlier* (AO) associated to Trend-Cycle:",
					knitr::combine_words(format(zoo::as.yearmon(outliers[[serie]]$ao_tc), "%B %Y")),
					"\n\n"
				)
			}

			if (!is.null(outliers[[serie]]$ls)) {
				cat(
					"\n\n- *Level shift* (LS):",
					knitr::combine_words(format(zoo::as.yearmon(outliers[[serie]]$ls), "%B %Y")),
					"\n\n"
				)
			}

		}

	}
}
```