### 2025-04-23 report/index.qmd

---
title: "Liste des publications"
format: 
  html:
    toc: true
    number-sections: true
    embed-resources: true
    number-depth: 2
    code-links:
      - text: AQLT/publishTC.wp
        icon: github
        href: https://github.com/AQLT/publishTC.wp
date: today
date-format: long
params:
  file: "../data/IPI.yml"
author:
  - name: Alain Quartier-la-Tente 
    affiliation: LEMNA
    orcid: 0000-0001-7890-3857
comments: 
    hypothesis:
      theme: clean
lang: fr
language:
 title-block-author-single: Auteur
---

```{r}
#| results: asis
#| echo: false
list_files <- list.files(pattern = "yml$", path = dirname(params$file), full.names = TRUE)
for (f in list_files) {
    data <- yaml::read_yaml(f)
    cat(sprintf("\n\n- [%s](%s.html) (%i series)",
        data$datasetname,
        data$dataset,
        length(data$series))
    )
}
```