### 2025-06-30 Article/template/Preambule.tex
linebreak inline code

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\med}{med}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\tr}{tr}
\usepackage[htt]{hyphenat}

\newcommand\1{{\mathds 1}}
\newcommand\ud{\,\mathrm{d}}
\newcommand{\transp}[1]{{}^t\!#1}
\newcommand{\bf}[1]{{\boldsymbol #1}}
\newcommand{\E}[1]{\mathbb{E}\left[#1\right]}
\newcommand{\V}[1]{\mathbb{V}\left[#1\right]}
\newcommand{\cov}[2]{\Cov\left(#1,#2\right)}

\usepackage{mathrsfs, dsfont}
\usepackage[style=authoryear,uniquename=false, uniquelist=false]{biblatex}
\DefineBibliographyStrings{french}{andothers={et\addabbrvspace alii}}